Study of the hadronization of charged pions

### 2016-10-07 chap6.tex
Updated the result part.

The results presented in this chapter are corrected for acceptance effects, 
but not for radiative ones. We use the following DIS selection cuts: 
$Q^2$~$>$~$1$~GeV$^2$/c$^2$, $W$~$>$~$2$ GeV and $y<0.85$. In order to ensure that 
factorization applies, we select $0.4$~$<$~$z$~$<$~$0.7$\footnote{This cut is not 
applied for plots as a function of $z$.}, this choice is based on experimental results from 
\cite{Asaturyan:2011mq}. Moreover, this cut also ensures that we measure the leading 
hadron and that we avoid the high $z$ region, which might be contaminated by 
the diffractive $\rho^0$ decay products. Finally, the cut $x_F > 0$ select the 
current fragmentation region.

@@

The Cronin effect is characterized by a large increase of the multiplicity 
ratio at high \pt ($\sim$1~GeV$^2$/c$^2$), but it is a controversial 
measurement in hadronization studies. Indeed, whereas SLAC measurements 
\cite{Osborne:1978ai} did not show such an effect, HERMES 
\cite{Airapetian:2007vu} measured a significant increase of the multiplicity 
ratio with \ptp. Our result (figure \ref{fig:RPt} (left)), integrated over 
all other variables, shows a pattern similar to the HERMES measurement. 
However, there is potential contributions from the target fragmentation (see 
chapter~\ref{chap:data}) and the Fermi motion (see chapter \ref{chap:pyqm}).

Fermi motion effects can be significantly reduced by modifying 
our usual observables. Indeed, using carbon for normalization, instead of deuterium, 
cancels most of the effects of Fermi motion. Moreover, acceptance 
effects (section~\ref{sec:accept}) are mostly canceled as well in such a ratio, 
reducing various systematic errors (including most of the normalization error). 
The figure \ref{fig:RPt} (right) represents the multiplicity ratio based on carbon. We 
observe an attenuation, at low \ptp, similar to the usual multiplicity ratio 
of iron. This can be understood from the radius of the nuclei $R_{Pb}-R_{C} 
\sim R_{Fe}$. Nevertheless, the observed enhancement with \pt is much more modest than in 
iron. At first sight the difference can be attributed to Fermi motion, which 
might affect the measurement based on deuterium. The target fragmentation 
effects might also lead to similar observation and more evidence are 
needed to confirm our interpretation.

In the figure \ref{fig:RPtMulti}, the multi-dimensionally binned multiplicity 
ratio are presented. HERMES found an important dependence of the Cronin effect 
with $z$ \cite{Airapetian:2007vu,Airapetian:2011jp} that we interpreted as a 
sign of an important target fragmentation contribution. The present result 
(figure~\ref{fig:RPtMulti} (left)) does not show this behavior, but it can be 
explained by our stricter cut on $z$ ($z>0.4$ whereas HERMES uses $z>0.2$), which 
leads to a smaller target fragmentation contamination. The second result, 
binned in \pt and $\nu$ (figure \ref{fig:RPtMulti} (right)), shows an important 
dependence of the Cronin effect with $\nu$. However, HERMES results did not 
reveal any similar $\nu$ variation. This is an evidence of the importance of 
the Fermi motion effect in our measurement. Indeed, in the 
chapter~\ref{chap:pyqm}, it was shown that the impact of Fermi motion gets 
smaller at higher energy.

@@

In conclusion, our results indicate that the HERMES results were affected by target 
fragmentation and that our results are affected by Fermi motion. These effects can 
be easily controlled by selecting carefully our observables and kinematic coverage.

@@

The HERMES collaboration clearly observed a rise of the multiplicity ratio 
between $\nu$ of 4 and 22~GeV. However, at first sight, in our results (figure 
\ref{fig:RNu} (left)), no dependence is observed, except for a 
slight increase in lead. The results on Fermi motion together 
with the figure \ref{fig:RNu} (right) permit to reach a coherent explanation. The 
Fermi motion seems to cancel the hadronization effect almost completely and the
systematic uncertainties linked with acceptance might wash out what remains of the effect. The 
multiplicity ratio based on carbon, however, offers a very neat slope consistent 
with what was observed on a much larger range by HERMES.

We also notice here that both pions give similar results in the figure 
\ref{fig:RNu} (right). This confirms our previous remark that the difference 
observed in the multiplicity ratio based on deuterium might come from the 
normalization uncertainty caused by the acceptance correction. Therefore, there is no 
clear difference between the charged pions for the integrated multiplicity 
ratios.

@@

The multiplicity ratio was observed to decrease with $z$ in HERMES data, 
whereas this feature was not as significant in other experiments. Indeed the 
nature of this behavior is questionable, the target fragmentation also reduce 
with $z$ and can mimic the effect. In figure \ref{fig:Rz} (left), as in 
HERMES, we see a clear slope even at values higher than 0.4, where target fragmentation
effects are expected to be small. However, the ratio lead to carbon (figure 
\ref{fig:Rz} (right)) shows a much flatter behavior in the region of interest 
(0.4 to 0.7). This is coherent with what we showed in chapter \ref{chap:pyqm}, i.e. the expected $z$ slope 
is enhanced by Fermi motion effects. Therefore, the situation seems to be 
similar to the Cronin effect one, where HERMES observation was enhanced by 
target fragmentation. The measurement using the carbon as basis is, therefore, 
more useful to isolate effects from the hadronization, but low $z$ behavior 
remains driven by the target region. Another strange feature of the data is 
the behavior at high $z$ (figure \ref{fig:Rz} (right)), where the two pions 
behave differently. However, we lack of solid theoretical grounds in this 
region to interpret this result.

@@

We can use our large statistics to get more information. As 
a dependence in $\nu$ is expected for the multiplicity ratio, it might be helpful to use a tight 
$\nu$ bin to plot the $Q^2$ dependence (figure \ref{fig:RQ2Detailed} (left)), 
and remove any coupling between the two variables.
We could also expect some effects from the Fermi motion, 
therefore, we show in figure \ref{fig:RQ2Detailed} (right) the multiplicity ratio of lead to 
carbon as a function of $Q^2$. The two results of figure \ref{fig:RQ2Detailed} 
show a slight increase with $Q^2$, but as for HERMES, no evidence is 
reached on this question. Our leverage on $Q^2$ appears to be too modest
to make a clear measurement. Future programs to explore this question are 
discussed in chapter~\ref{chap:future}.

@@

The $A$ dependence of the transverse momentum broadening, presented in 
figure~\ref{fig:PA}, is an important result. The large statistics coupled with the 
large coverage in $A$ available to the CLAS experiment give precise indication on \dpt 
as a function of $A^{1/3}$. Compared to HERMES \cite{Airapetian:2009jy}, we
find smaller \dptp. This is coherent with theory, which predicts larger effects
at larger energies. However, calculations of parton energy loss from BDMPS 
\cite{Baier:1996sk} show that \pt should depend on the nucleus radii
(see chapter \ref{chap:theory}). However, this feature does not appear in our 
result. Of course, because of the low energy of the CLAS experiment, one might 
discuss the possibility to make a direct comparison with BDMPS calculations. Nonetheless, 
this result shows an unexpected pattern that remains to be explained. One 
possibility is that, as proposed for the multiplicity ratio as a function of 
$A^{1/3}$, the production time occurs inside the nuclei. In this case, the 
colored parton would not interact with the whole nuclei and therefore changing 
its size would lead to limited effect.

@@

The $z$ dependence of \dpt has some importance for parton energy loss
models, because it is linked to the assumptions needed to extrapolate the hadronic \pt
to the partonic one. Moreover, usually, in parton energy loss models, \dpt
goes to zero at the highest $z$. Our results (figure \ref{fig:PZ}) do
not show such a pattern. However, our previous results showed the importance of 
Fermi motion in our kinematics, which is expected to enhance \dpt at very high
$z$. This is confirmed by the flattening observed for the carbon based \dptp, but
the distribution is not clearly going down either at high $z$. This might be due to the size of the 
error bars and the potential contamination from diffractive $\rho^0$ production, both 
preventing us from any definitive statement on this feature of the data.

\begin{figure}[tbp]
\centering
\includegraphics[width=7.4cm] {chap6-fig/F_PvZ.png} 
\includegraphics[width=7.4cm] {chap6-fig/F_PvZ_PbC.png} 
\caption {\dpt as a function of $z$. Left panel is the usual
observable, right panel shows multiplicity ratio of lead normalized to carbon. Normalization uncertainties are not shown.}
\label{fig:PZ}
\end{figure}

>>>>

The results presented in this chapter are a selection of the observables 
that we want to use in a first publication, eventually a long paper would
include all results as function of all variables. We use the following DIS selection cuts: 
$Q^2$~$>$~$1$~GeV$^2$/c$^2$, $W$~$>$~$2$ GeV and $y<0.85$. In order to ensure that 
factorization applies, we select $0.4$~$<$~$z$~$<$~$0.7$\footnote{This cut is not 
applied for plots as a function of $z$.}, this choice is based on experimental results from 
\cite{Asaturyan:2011mq}. Moreover, this cut also ensures that we measure the leading 
hadron and that we avoid the high $z$ region, which might be contaminated by 
the decay products of the diffractive production of $\rho^0$. Finally, the cut 
$x_F > 0$ select the forward fragmentation region.

@@

The Cronin effect is characterized by a large increase of the multiplicity 
ratio at high \pt ($\sim$1~GeV$^2$/c$^2$), but it is a controversial 
measurement in hadronization studies. Indeed, whereas SLAC measurements 
\cite{Osborne:1978ai} did not show such an effect, HERMES 
\cite{Airapetian:2007vu} measured a significant increase of the multiplicity 
ratio with \ptp. Our result (figure \ref{fig:RPt} (left)), integrated over 
all other variables, shows a pattern similar to the HERMES measurement. 
However, there is potential contributions from the target fragmentation and 
the Fermi motion.

Fermi motion effects can be significantly reduced by modifying 
our usual observables. Indeed, using carbon for normalization, instead of deuterium, 
cancels most of the effects of Fermi motion. Moreover, acceptance 
effects (section~\ref{sec:accept}) are mostly canceled as well in such a ratio, 
reducing the associated systematic errors (including most of the normalization error). 
The figure \ref{fig:RPt} (right) represents the multiplicity ratio based on carbon. We 
observe an attenuation, at low \ptp, similar to the usual multiplicity ratio 
of iron. This can be understood from the radius of the nuclei $R_{Pb}-R_{C} 
\sim R_{Fe}$. Nevertheless, the observed enhancement with \pt is much more modest than in 
iron. At first sight the difference can be attributed to Fermi motion, which 
might affect the measurement based on deuterium.

In the figure \ref{fig:RPtMulti}, the multi-dimensionally binned multiplicity 
ratio are presented. HERMES found an important dependence of the Cronin effect 
with $z$ \cite{Airapetian:2007vu,Airapetian:2011jp} that can be interpreted as a 
sign of an important target fragmentation contribution. The present result 
(figure~\ref{fig:RPtMulti} (left)) does not show this behavior, but it can be 
explained by our stricter cut on $z$ ($z>0.4$ whereas HERMES uses $z>0.2$), which 
leads to a smaller target fragmentation contamination. The second result, 
binned in \pt and $\nu$ (figure \ref{fig:RPtMulti} (right)), shows an important 
dependence of the Cronin effect with $\nu$. However, HERMES results did not 
reveal any similar $\nu$ variation. This is an evidence of the importance of 
the Fermi motion effect in our measurement.

@@

In conclusion, our results indicate that the HERMES results were affected by target 
fragmentation and that our results are affected by Fermi motion. These effects can 
be easily controlled by selecting carefully our observables. In this regard, the
Pb/C ratio seems to give cleaner results than Fe/D.

@@

The HERMES collaboration clearly observed a rise of the multiplicity ratio 
between $\nu$ of 4 and 22~GeV. However, at first sight, in our results (figure 
\ref{fig:RNu} (left)), no dependence is observed, except for a 
slight increase in lead.
However, Fermi motion could cancel the hadronization effect almost completely and the
systematic uncertainties linked with acceptance might wash out what remains of the effect. The 
multiplicity ratio based on carbon, however, offers a cleaner result with a slope consistent 
with what was observed on a much larger range by HERMES.

We also notice here that both pions give similar results in the figure 
\ref{fig:RNu} (right). This confirms our previous remark that the difference 
observed in the multiplicity ratio based on deuterium might come from the 
normalization uncertainty caused by the acceptance correction. There is no 
clear difference between the charged pions for the integrated multiplicity 
ratios as a function of $\nu$.

@@

The multiplicity ratio was observed to decrease with $z$ in HERMES data, 
whereas this feature was not as significant in other experiments. Indeed the 
nature of this behavior is questionable, the target fragmentation also reduce 
with $z$ and can mimic the effect. In figure \ref{fig:Rz} (left), as in 
HERMES, we see a clear slope even at values higher than 0.4, where target fragmentation
effects are expected to be small. However, the ratio lead to carbon (figure 
\ref{fig:Rz} (right)) shows a much flatter behavior in the region of interest 
(0.4 to 0.7). Therefore, the situation seems to be 
similar to the Cronin effect one, where HERMES observation was enhanced by 
target fragmentation. The measurement using the carbon as basis is, therefore, 
more useful to isolate effects from the hadronization, but low $z$ behavior 
remains driven by the target region. Another strange feature of the data is 
the behavior at high $z$ (figure \ref{fig:Rz} (right)), where the two pions 
behave differently. However, we lack of solid theoretical grounds in this 
region to interpret this result.

@@

We can use our large statistics to get more information. As 
a dependence in $\nu$ is expected for the multiplicity ratio, it might be helpful to use a tight 
$\nu$ bin to plot the $Q^2$ dependence (figure \ref{fig:RQ2Detailed} (left)), 
and remove any coupling between the two variables.
We could also expect some effects from the Fermi motion, 
therefore, we show in figure \ref{fig:RQ2Detailed} (right) the multiplicity ratio of lead to 
carbon as a function of $Q^2$. The two results of figure \ref{fig:RQ2Detailed} 
show a slight increase with $Q^2$, but as for HERMES, no evidence is 
reached on this question. Our leverage on $Q^2$ appears to be too modest
to make a clear measurement.

@@

The $A$ dependence of the transverse momentum broadening, presented in 
figure~\ref{fig:PA}, is an important result. The large statistics coupled with the 
large coverage in $A$ available to the CLAS experiment give precise indication on \dpt 
as a function of $A^{1/3}$. Compared to HERMES \cite{Airapetian:2009jy}, we
find smaller \dptp. This is coherent with theory, which predicts larger effects
at larger energies. However, calculations of parton energy loss from BDMPS 
\cite{Baier:1996sk} show that \pt should depend on the nucleus radii. However, 
this feature does not appear in our 
result. Of course, because of the low energy of the CLAS experiment, one might 
discuss the possibility to make a direct comparison with BDMPS calculations. Nonetheless, 
this result shows an unexpected pattern that remains to be explained. One 
possibility is that, as proposed for the multiplicity ratio as a function of 
$A^{1/3}$, the production time occurs inside the nuclei. In this case, the 
colored parton would not interact with the whole nuclei and limit the nuclear 
effect.

@@
